---
title: "final-review-data"
author: "JP Flores"
date: "2023-07-27"
output: html_document
---

These data were collected during two research cruises in the North Atlantic in May and October 2003. During these two months the research vessel visited multiple stations (areas) and marine mammal observers recorded the number of whales (who doesn’t love whales!) at each of these stations. The time the vessel spent at each station was also recorded along with other site specific variables such as the latitude and longitude, water depth and gradient of the seabed. The researchers also recorded the ambient level of sub-surface noise with a hydrophone and categorised this variable into ‘low’, ‘medium’ or ‘high’. The structure of these data is known as a rectangular dataset (aka ‘tidy’ data by the cool kids) with no missing cells. Each row is an individual observation and each column a separate variable. The variable names are contained in the first row of the dataset (aka a header).

```{r}
## 1) Import the data


## 2) How can you tell the dimensions of the data? How many rows? How many columns?

```


What type of variables are month and water.noise?
```{r}

```


You can get another useful summary of your dataframe by using the summary() function. This will provide you with some useful summary statistics for each variable. Notice how the type of output depends on whether the variable is a factor or a number. Another useful feature of the summary() function is that it will also count the number of missing values in each variable. Which variables have missing values and how many?

```{r}

```


Summarising and manipulating dataframes is a key skill to acquire when learning R. Although there are many ways to do this, we will concentrate on using the square bracket [ ] notation which you used previously with vectors. The key thing to remember when using [ ] with dataframes is that dataframes have two dimensions (think rows and columns) so you always need to specify which rows and which columns you want inside the [ ] (see Section 3.4.1 and this video for some additional background information and a few examples). Let’s practice.

a) Extract all the elements of the first 10 rows and the first 4 columns of the whale dataframe and assign to a new variable called whale.sub.
```{r}

```


Next, extract all observations (remember - rows) from the whale dataframe and the columns month, water.noise and number.whales and assign to a variable called whale.num.
```{r}

```


Now, extract the first 50 rows and all columns form the original dataframe and assign to a variable whale.may (there’s a better way to do this with conditional statements - see below).
```{r}

```


Finally, extract all rows except the first 10 rows and all columns except the last column. Remember, for some of these questions you can specify the columns you want either by position or by name. Practice both ways. Do you have a preference? If so why?
```{r}

```


In addition to extracting rows and columns from your dataframe by position you can also use conditional statements to select particular rows based on some logical criteria. This is very useful but takes a bit of practice to get used to (see Section 3.4.2 for an introduction). Extract rows from your dataframe (all columns by default) based on the following criteria (note: you will need to assign the results of these statements to appropriately named variables, I’ll leave it up to you to use informative names!):

a) at depths greater than 1200 m
```{r}

```

b) gradient steeper than 200 degrees

```{r}

```

c) water noise level of ‘low’
```{r}

```

d) water.noise level of ‘high’ in the month of ‘May’

```{r}

```

e) month of ‘October’, water noise level of ‘low’ and gradient greater than the median value of gradient (132)
```{r}

```

f) all observations from between latitudes 60.0 and 61.0 and longitudes -6.0 and -4.0
```{r}

```

g) all rows that do not have a water noise level of medium

```{r}

```


10. A really neat feature of extracting rows based on conditional statements is that you can include R functions within the statement itself. To practice this, modify your answer to the gradient question in Q9 to use the median() function rather than hard coding the value 132.
```{r}

```



11. However, when using functions in conditional statements you need to be careful. For example, write some code to extract all rows from the dataframe whale with depths greater than 1500 m and with a greater number of whales spotted than average (hint: use the mean() function in your conditional statement). Can you see a problem with the output? Discuss the cause of this problem with an instructor and explore possible solutions.

```{r}

```


12. Although you have concentrated on using the square bracket [ ] notation to extract rows and columns from your dataframe, there are of course many other approaches. One such approach is to use the subset() function (see ?subset or search for the subset function in the Introduction to R book to find more information). Use the subset() function to extract all rows in ‘May’ with a time at station less than 1000 minutes and a depth greater than 1000 m. Also use subset() to extract data collected in ‘October’ from latitudes greater than 61 degrees but only include the columns month, latitude, longitude and number.whales.
```{r}

```



13. Another useful way to manipulated your dataframes is to sort the rows based on the value of a variable (or combinations of variables). Rather counter-intuitively you should use the order() function to sort your dataframes, not the sort() function (see Section 3.4.3 of the Introduction to R book for an explanation). Ordering dataframes uses the same logic you practised in Q14 in Exercise 2. Let’s practice with a straight forward example. Use the order() function to sort all rows in the whale dataframe in ascending order of depth (shallowest to deepest). Store this sorted dataframe in a variable called whale.depth.sort.
```{r}

```




14. Now for something a little more complicated. Sort all rows in the whale dataframe by ascending order of depth within each level of water noise. The trick here is to remember that you can order by more than one variable when using the order() function (see Section 3.4.3 again). Don’t forget to assign your sorted dataframe to a new variable with a sensible name. Repeat the previous ordering but this time order by descending order of depth within each level of water noise.

```{r}

```


18. Ok, we have spent quite a bit of time (and energy) learning how to import and manipulate dataframes. The last thing we need to cover is how to export dataframes from R to an external file (see Section 3.6 of the book for more details). Let’s say you want to export the dataframe whale.num you created previously (see Q8) to a file called ‘whale_num.txt’ in your output directory which you created in Exercise 1. To do this you will need to use the write.table() function. You want to include the the variable names in the first row of the file, but you don’t want to include the row names. Also, make sure the file is a tab delimited file. Once you have create your file, try to open it in Microsoft Excel (or open source equivalent).
```{r}

```







