---
title: "structures"
author: "JP Flores"
date: "2023-07-11"
output: html_document
---
### Exercises

These exercises are to help you solidify and expand on the information given above. We intentionally added some concepts that were not covered above, and hope that you will take a few minutes to think through what is happening and how R is interpreting the code. 

1. Create a data frame with three columns: "col1" with the letters a to f, "col2" with the numbers 1 to 6, and "col3" with the alternating TRUE and FALSE values. Store the data frame as "mydf."

```{r}

```

2. Coerce `mydf` from exercise 1 to a matrix using the `as.matrix` function. Predict how the data will change. What class will the data be? Will there be column names? Row names?

```{r}

```


3. Subsetting can also be done with logical vectors the same length of the object or dimension you would like to subset. For example, if you have a list `l` with three elements, `l[c(TRUE, FALSE, TRUE)]` would return a list with elements one and three from `l`. Figure out how to use `mydf` to subset only to rows where 'col3' is `TRUE`. (You should attempt to think of a solution that only requires one line of code. Hint: you can pass elements of an obect to itself.)

```{r}

```


4. You can stack subsetting operators next to each other. Using `l2` from above, select the 7 from 'vec.' Now try to select the last two rows from 'mat'. (Again, you should attempt to think of a solution that only requires one line of code for each selection.)

```{r}

```


5. Recall how you added an element to `l2` using the `$` operator. Again, a data frame is just a special list. Add a column to `mydf` called "col4" with a vector of your choice. 

```{r}

```

